---
title: "a3_task2_dani_sclafani"
author: "Danielle Sclafani"
date: "2/17/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster) 
library(tidyverse)
library(here)
library(sf)
library(janitor)
library(fasterize)
library(rnaturalearth)
library(sp)
library(rnaturalearthdata)
library(rgeos)

```

## Background
The abundance of individual cetacean species along the California coast is approximated based on the relative environmental suitability, including species preferences for water temperature, depth, salinity, and distance to land. The probability of occurrence of 35 cetacean species are included in the AquaMaps dataset used in this report. The data is used to make a rasterized map, showing the probable species abundance based on a probability threshold of at least 0.6. The extent of the rasters is 125째 W to 115째W (which R considers -125 to -115) and 32째N to 38째N.  

### Data Wrangling Part 1
```{r, message = FALSE, warning = FALSE}
#loading in the data

total_cetaceans <-list.files(path = here("ca_cetaceans"), full.names = TRUE)

ceta_stack <- raster::stack(total_cetaceans)

ceta_stack_df <- raster::rasterToPoints(ceta_stack) %>% 
  as.data.frame() 

ceta_df <- ceta_stack_df %>% 
  mutate_all(~replace(., is.na(.),0)) %>% 
  mutate(
    Balaenoptera_acutorostrata = case_when(Balaenoptera_acutorostrata > 0.6 ~ 1),
    Balaenoptera_borealis = case_when(Balaenoptera_borealis > 0.6 ~ 1),
    Balaenoptera_brydei = case_when(Balaenoptera_brydei >0.6 ~ 1),
    Balaenoptera_edeni = case_when(Balaenoptera_edeni > 0.6 ~ 1),
    Balaenoptera_musculus = case_when(Balaenoptera_musculus > 0.6 ~1),
    Balaenoptera_physalus = case_when(Balaenoptera_physalus > 0.6 ~1),
    Berardius_bairdii = case_when(Berardius_bairdii > 0.6 ~1),
    Delphinus_capensis = case_when(Delphinus_capensis > 0.6 ~1),
    Delphinus_delphis = case_when(Delphinus_delphis > 0.6 ~1),
    Eschrichtius_robustus = case_when( Eschrichtius_robustus > 0.6 ~1),
    Eubalaena_japonica = case_when(Eubalaena_japonica> 0.6 ~1),
    Globicephala_macrorhynchus = case_when(Globicephala_macrorhynchus > 0.6 ~1),
    Grampus_griseus = case_when(Grampus_griseus > 0.6 ~1),
    Indopacetus_pacificus = case_when(Indopacetus_pacificus > 0.6 ~1),
    Kogia_breviceps = case_when(Kogia_breviceps > 0.6 ~1),
    Kogia_sima = case_when(Kogia_sima > 0.6 ~ 1),
    Lagenodelphis_hosei = case_when(Lagenodelphis_hosei > 0.6 ~1),
    Lagenorhynchus_obliquidens = case_when(Lagenorhynchus_obliquidens > 0.6~ 1),
    Lissodelphis_borealis = case_when(Lissodelphis_borealis > 0.6 ~1),
    Megaptera_novaeangliae = case_when(Megaptera_novaeangliae > 0.6 ~1),
    Mesoplodon_carlhubbsi = case_when(Mesoplodon_carlhubbsi > 0.6 ~1),
    Mesoplodon_densirostris = case_when(Mesoplodon_densirostris > 0.6 ~1),
    Mesoplodon_ginkgodens = case_when(Mesoplodon_ginkgodens > 0.6 ~ 1),
    Mesoplodon_perrini = case_when(Mesoplodon_perrini > 0.6 ~1),
    Mesoplodon_stejnegeri = case_when(Mesoplodon_stejnegeri > 0.6 ~ 1),
    Orcinus_orca = case_when(Orcinus_orca > 0.6 ~1),
    Phocoena_phocoena = case_when(Phocoena_phocoena > 0.6 ~1),
    Phocoenoides_dalli = case_when(Phocoenoides_dalli > 0.6 ~1),
    Physeter_macrocephalus = case_when(Physeter_macrocephalus > 0.6 ~1),
    Pseudorca_crassidens = case_when(Pseudorca_crassidens > 0.6 ~1),
    Stenella_attenuata = case_when(Stenella_attenuata > 0.6 ~1),
    Stenella_coeruleoalba = case_when(Stenella_coeruleoalba > 0.6 ~1),
    Steno_bredanensis = case_when(Steno_bredanensis > 0.6 ~1),
    Tursiops_truncatus = case_when(Tursiops_truncatus > 0.6 ~ 1),
    Ziphius_cavirostris = case_when(Ziphius_cavirostris > 0.6 ~1)
  )%>% 
  mutate_all(~replace(., is.na(.),0)) %>% 
  clean_names()
  
ceta_df$total_prob <- rowSums(ceta_df[,3:37], na.rm = TRUE)

xyz_ceta_df <- ceta_df %>% 
  select(x, y, total_prob)

ggplot(data = xyz_ceta_df, aes(x =x, y=y, fill = total_prob)) + geom_raster()

```


### Data Wrangling Part 2 - proved more successful
```{r, warning = FALSE, message=FALSE}
# trying with something else, still need to establish a threshold
test_total <-list.files(path = here("ca_cetaceans"), full.names = TRUE)
 
stack_test <- raster::stack(test_total)

plot(stack_test)
```

### Creating the threshold
```{r, message = FALSE, warning= FALSE}
#establishing the threshold

is_species <-function(x, thresh = 0.6) {
  y <- ifelse (x >= thresh, 1, 0)
  return(y)
}

species <- calc(stack_test, fun = is_species)

plot(species) #plotting again to see how the threshold changed things
```



### Converting the raster data to a data frame for ggplot
```{r}
#converting data frame for ggplot
species_df <-raster::rasterToPoints(species) %>% 
  as.data.frame() %>% 
  mutate_all(~replace(., is.na(.), 0))

species_df$total_prob <- rowSums(species_df[,3:37], na.rm = TRUE) #this summed all the probabilities that were above the threshold to get the total probability of species abundance 


#did not need a stack_raster_df
#stack_raster_df <-raster::rasterToPoints(stack_test) %>% 
  #as.data.frame() 

#ggplot(data = stack_raster_df, aes(x=x, y=y)) + geom_raster()+
#  geom_raster(data = species_df, fill = 'blue') +
#  coord_sf(expand = 0)+
 # scale_fill_gradient(low = 'lightblue', high= 'navy')+
# theme_void()+
 # theme(panel.background = element_rect(fill = 'slateblue4'))

```


### Loading in the Map form `rnaturalearth` Package
```{r}
#loading in the map from rnatural earth
map <- ne_countries(scale = "medium", returnclass = "sf")

ggplot(data = map) +
  geom_sf() +
  coord_sf(xlim = c(-125, -115), ylim = c(32,38), expand = FALSE)
```


### Graphing the Map and the Raster Together
```{r}

ggplot()+ 
  geom_raster(data = species_df, aes( x=x, y =y, fill = total_prob))+
  scale_fill_gradient(low = 'yellow', high= 'red')+
  theme_void()+
  geom_sf(data = map, fill = "lightgreen")+
  coord_sf(xlim = c(-125, -115), ylim = c(32,38), expand = FALSE) #set coordinates to the same as the raster so they match up

```


Data Citation:
Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org
