---
title: "a3_task2_dani_sclafani"
author: "Danielle Sclafani"
date: "2/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster) 
library(tidyverse)
library(here)
library(sf)
library(janitor)
library(fasterize)
library(rnaturalearth)
library(sp)
library(rnaturalearthdata)
library(rgeos)
```

showing the probability of occurrence (based on relative environmental suitability, including species preferences for water temperature, depth, salinity, and distance to land) of 35 cetacean species that can be found in the California Bight.  The extent of the rasters is 125째 W to 115째W (which R considers -125 to -115) and 32째N to 38째N.  


```{r}
#loading in the data

total_cetaceans <-list.files(path = here("ca_cetaceans"), full.names = TRUE)

ceta_stack <- raster::stack(total_cetaceans)

ceta_stack_df <- raster::rasterToPoints(ceta_stack) %>% 
  as.data.frame() 

ceta_df <- ceta_stack_df %>% 
  mutate_all(~replace(., is.na(.),0)) %>% 
  mutate(
    Balaenoptera_acutorostrata = case_when(Balaenoptera_acutorostrata > 0.6 ~ 1),
    Balaenoptera_borealis = case_when(Balaenoptera_borealis > 0.6 ~ 1),
    Balaenoptera_brydei = case_when(Balaenoptera_brydei >0.6 ~ 1),
    Balaenoptera_edeni = case_when(Balaenoptera_edeni > 0.6 ~ 1),
    Balaenoptera_musculus = case_when(Balaenoptera_musculus > 0.6 ~1),
    Balaenoptera_physalus = case_when(Balaenoptera_physalus > 0.6 ~1),
    Berardius_bairdii = case_when(Berardius_bairdii > 0.6 ~1),
    Delphinus_capensis = case_when(Delphinus_capensis > 0.6 ~1),
    Delphinus_delphis = case_when(Delphinus_delphis > 0.6 ~1),
    Eschrichtius_robustus = case_when( Eschrichtius_robustus > 0.6 ~1),
    Eubalaena_japonica = case_when(Eubalaena_japonica> 0.6 ~1),
    Globicephala_macrorhynchus = case_when(Globicephala_macrorhynchus > 0.6 ~1),
    Grampus_griseus = case_when(Grampus_griseus > 0.6 ~1),
    Indopacetus_pacificus = case_when(Indopacetus_pacificus > 0.6 ~1),
    Kogia_breviceps = case_when(Kogia_breviceps > 0.6 ~1),
    Kogia_sima = case_when(Kogia_sima > 0.6 ~ 1),
    Lagenodelphis_hosei = case_when(Lagenodelphis_hosei > 0.6 ~1),
    Lagenorhynchus_obliquidens = case_when(Lagenorhynchus_obliquidens > 0.6~ 1),
    Lissodelphis_borealis = case_when(Lissodelphis_borealis > 0.6 ~1),
    Megaptera_novaeangliae = case_when(Megaptera_novaeangliae > 0.6 ~1),
    Mesoplodon_carlhubbsi = case_when(Mesoplodon_carlhubbsi > 0.6 ~1),
    Mesoplodon_densirostris = case_when(Mesoplodon_densirostris > 0.6 ~1),
    Mesoplodon_ginkgodens = case_when(Mesoplodon_ginkgodens > 0.6 ~ 1),
    Mesoplodon_perrini = case_when(Mesoplodon_perrini > 0.6 ~1),
    Mesoplodon_stejnegeri = case_when(Mesoplodon_stejnegeri > 0.6 ~ 1),
    Orcinus_orca = case_when(Orcinus_orca > 0.6 ~1),
    Phocoena_phocoena = case_when(Phocoena_phocoena > 0.6 ~1),
    Phocoenoides_dalli = case_when(Phocoenoides_dalli > 0.6 ~1),
    Physeter_macrocephalus = case_when(Physeter_macrocephalus > 0.6 ~1),
    Pseudorca_crassidens = case_when(Pseudorca_crassidens > 0.6 ~1),
    Stenella_attenuata = case_when(Stenella_attenuata > 0.6 ~1),
    Stenella_coeruleoalba = case_when(Stenella_coeruleoalba > 0.6 ~1),
    Steno_bredanensis = case_when(Steno_bredanensis > 0.6 ~1),
    Tursiops_truncatus = case_when(Tursiops_truncatus > 0.6 ~ 1),
    Ziphius_cavirostris = case_when(Ziphius_cavirostris > 0.6 ~1)
  )%>% 
  mutate_all(~replace(., is.na(.),0)) %>% 
  clean_names()
  
ceta_df$total_prob <- rowSums(ceta_df[,3:37], na.rm = TRUE)

xyz_ceta_df <- ceta_df %>% 
  select(x, y, total_prob)

ggplot(data = xyz_ceta_df, aes(x =x, y=y, fill = total_prob)) + geom_raster()

```


# trying with something else, still need to establish a threshold
```{r}
test_total <-list.files(path = here("ca_cetaceans"), full.names = TRUE)
 
stack_test <- raster::stack(test_total)

plot(stack_test)
```


```{r}
#establishing the threshold

is_species <-function(x, thresh = 0.6) {
  y <- ifelse (x >= thresh, 1, 0)
  return(y)
}

species <- calc(stack_test, fun = is_species)

plot(species)
```




```{r}
#converting data frame for ggplot
species_df <-raster::rasterToPoints(species) %>% 
  as.data.frame() %>% 
  mutate_all(~replace(., is.na(.), 0))

species_df$total_prob <- rowSums(species_df[,3:37], na.rm = TRUE)


#did not need a stack_raster_df
#stack_raster_df <-raster::rasterToPoints(stack_test) %>% 
  as.data.frame() 

#ggplot(data = stack_raster_df, aes(x=x, y=y)) + geom_raster()+
#  geom_raster(data = species_df, fill = 'blue') +
#  coord_sf(expand = 0)+
 # scale_fill_gradient(low = 'lightblue', high= 'navy')+
# theme_void()+
 # theme(panel.background = element_rect(fill = 'slateblue4'))

```


## loading in the map form rnatural earth
```{r}
#loading in the map from rnatural earth
map <- ne_countries(scale = "medium", returnclass = "sf")

ggplot(data = map) +
  geom_sf() +
  coord_sf(xlim = c(-125, -115), ylim = c(32,38), expand = FALSE)
```

## graphing the map and the raster together
```{r}
ggplot()+ 
  geom_raster(data = species_df, aes( x=x, y =y, fill = total_prob))+
  scale_fill_gradient(low = 'yellow', high= 'red')+
  theme_void()+
  geom_sf(data = map, fill = "lightgreen")+
  coord_sf(xlim = c(-125, -115), ylim = c(32,38), expand = FALSE)


```



```{r}

# trying to make res smaller, think we can't since beginning res was already so big.
res(species) #res = 0.5, 0.5

species_new_res <- aggregate(species, fact = 0./res(species))
```


### trying to make map earth topographic

```{r}
map_1 <- ne_download(scale = 110, category = "physical", type = 'land')


ggplot()+ 
  geom_raster(data = species_df, aes( x=x, y =y, fill = total_prob))+
  scale_fill_gradient(low = 'yellow', high= 'red')+
  theme_void()+
  geom_sf(data = coastline110)+
  coord_sf(xlim = c(-125, -115), ylim = c(32,38), expand = FALSE)

coastline110 <- ne_download(scale = 110, type = 'land', category = 'physical', returnclass = "sf")
```



